---
title: "p8105_hw2_yx2710"
output: github_document
---

```{r setup}
library(tidyverse)
library(readxl)
```

Problem 1 - nyc  transit data

```{r}
nyc_transit_df = read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv")%>%
  janitor::clean_names()%>%
  select(line, station_name, station_latitude, station_longitude, starts_with("route"), entry, vending, entrance_type, ada)%>%
  mutate(entry = as.logical(ifelse(entry == "YES", "TRUE", "FALSE")))
```
The variables that nyc transit dataset contains are `r names(nyc_transit_df)` and it has `r ncol(nyc_transit_df)` columns and `r nrow(nyc_transit_df)` rows.

Q:How many distinct stations are there?
```{r}
station = distinct(nyc_transit_df, station_name, line, .keep_all = TRUE)
nrow(station)
```
A:There are `r nrow(station)` distinct stations.

Q:How many stations are ADA compliant?
```{r}
station_ada = filter(nyc_transit_df, ada == "TRUE")
nrow(station_ada)
```
A: There are `r nrow(station_ada)` stations that are ADA compliant.

Q:What proportion of station entrances / exits without vending allow entrance?
```{r}
proportion_e_nov = nrow(filter(nyc_transit_df, entry == "TRUE", vending == "NO"))/nrow(filter(nyc_transit_df, vending == "NO"))
```
A: The proportion of station entrances / exits without vending allow entrance is `r proportion_e_nov`.

Reformat data so that route number and route name are distinct variables. How many distinct stations serve the A train? Of the stations that serve the A train, how many are ADA compliant?
```{r}
nyc_transit_re_df = 
  nyc_transit_df%>%
  mutate(
    route8 = as.character(route8),
    route9 = as.character(route9),
    route10 = as.character(route10),
    route11 = as.character(route11))%>%
  pivot_longer(route1:route11, 
               names_to = "route",
               values_to = "route_number")

station_re = distinct(nyc_transit_re_df, station_name, line, .keep_all = TRUE)

station_A = filter(station_re, route_number == "A" )
nrow(station_A)

station_A_ada = filter(station_A, ada == "TRUE")
nrow(station_A_ada)
```
There are `r nrow(station_A)` distinct stations serve the A train.
Of the stations that serve the A train, `r nrow(station_A_ada)` are ADA compliant.



Problem2 - trash
Mr. Trash Wheel sheet:
```{r}
mr_trash_df = 
  read_excel("./data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", sheet = "Mr. Trash Wheel")%>%
  janitor::clean_names()%>%
  select(dumpster:homes_powered) %>%
  drop_na(dumpster)%>%
  filter(dumpster != "Grand Total")%>%
  mutate(sports_balls = as.integer(round(sports_balls)))

```

Professor Trash Wheel:
```{r}
prof_trash_df = 
  read_excel("./data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", sheet = "Professor Trash Wheel")%>%
  janitor::clean_names()%>%
  select(dumpster:homes_powered) %>%
  drop_na(dumpster)%>%
  mutate(sports_balls = as.integer(round(sports_balls)))
```

Combine two dataset
```{r}
mr_trash_bind_df = mr_trash_df %>%
  mutate(cleaner = "Mr. Trash") %>%
  mutate(dumpster = as.numeric(dumpster))

prof_trash_bind_df = mutate(prof_trash_df, cleaner = "Professor Trash")
class(prof_trash_bind_df$dumpster)

trash_combine_df = 
  bind_rows(mr_trash_bind_df, prof_trash_bind_df)
```

Problem3 - fivethirtyeight
clean the data in pols-month.csv
```{r}
pols_month_df = 
  read_csv("./data/fivethirtyeight_datasets/pols-month.csv")%>%
  janitor::clean_names()%>%
  separate(col = mon, into = c("year", "month", "day"), sep = "-")%>%
  mutate(month = as.numeric(month), 
         month = month.abb[month],
         president = ifelse(prez_gop == 0, "dem", "gop"))%>%
  select(-prez_dem, -prez_gop, -day)
```

clean the data in snp.csv 
```{r}
snp_df = 
  read_csv("./data/fivethirtyeight_datasets/snp.csv")%>%
  janitor::clean_names()%>%
  mutate(date = lubridate::parse_date_time2(date, "mdy", cutoff_2000 = 49))%>%
  separate(col = date, into = c("year", "month", "day"), sep = "-")%>%
  mutate(month = as.numeric(month), 
         month = month.abb[month])%>%
  select(-day)
```

tidy the unemployment data
```{r}
unemp_df = 
  read_csv("./data/fivethirtyeight_datasets/unemployment.csv")%>%
  janitor::clean_names()%>%
  pivot_longer(
    jan:dec,
    names_to = "month",
    values_to = "unemployement"
  )
```








